---
title: "Dog Project"
author: "Allyson Hafezi, Fiona Davoudi, and Jenna LastName"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load the libraries 
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(naniar)
library("ggthemes")
library("RColorBrewer")
library("paletteer")
``` 

## Organization of Project

Todo List

1. Clean Data
    - make nice names with underscores 
    - remove $ and , and % from variables in dog_rankings 
    - change chr to num variables when needed
    - clean shelter_dogs?/maybe pivot some columns in either dataset 
    
2. Introduction - what is our research topic? List ideas below:
    - Making an App to choose a dog breed (dogs_ranking) and what are their traits, health situations. Combine with real shelter data on how many of that breed are typically available. 

3. Methods - how will we analyze
    - make quiz shiny app so people can figure out what dog/dog breed fits their needs
    - create graphs showing notatble characteristics associated with popularity or that correlate with each other

4. Visuals - graphs, what are our key variables?

5. Conclusion - clear message

6. Organize Repository (going off the example given, but we can refine)
    - datasets folder
    - background info folder - need to check if we have any
    - shiny app folder
    - exploring data
    - graphs

## Read in the Data

"This data is a compilation of information on dogs who were available for adoption on December 12, 2019 in the Hungarian Database of Homeless Pets. In total, there were 2,937 dogs in the database. It contains information on dogs' names, breed, color, age, sex, the date they were found, and some characteristics of their personalities."
[shelter_dogs](https://www.kaggle.com/datasets/jmolitoris/adoptable-dogs)
```{r}
shelter_dogs <- read_csv("datasets/ShelterDogs.csv") %>% clean_names()
```

"This dataset provides various attributes related to different dog breeds. Each row represents a specific breed and includes information on several characteristics, costs, and suitability aspects."
[dog_rankings](https://www.kaggle.com/datasets/jainaru/dog-breeds-ranking-best-to-worst)
```{r}
dog_rankings <- read_csv("datasets/dogs-ranking-dataset.csv") %>% clean_names
```

## Evaluate Data

```{r}
glimpse(shelter_dogs)
```

```{r}
glimpse(dog_rankings)
```

## Data Cleaning - Fiona
```{r}
#change name of size_5 and size_9 column in dog_rankings
dog_rankings_c <- dog_rankings %>% 
  rename(size_score=size_5) %>% 
  rename(size_description=size_9)
```

```{r}
#remove $ or , or % from entries
dog_rankings_c <- dog_rankings_c %>% 
  mutate(lifetime_cost = str_remove(lifetime_cost, ",")) %>% 
  mutate(lifetime_cost = str_remove(lifetime_cost, "\\$")) %>% 
  mutate(intelligence_percent = str_remove(intelligence_percent, "%")) %>% 
  mutate(purchase_price = str_remove(purchase_price, "\\$")) %>% 
  mutate(purchase_price = str_remove(purchase_price, ",")) %>% 
  mutate(food_costs_per_year = str_remove(food_costs_per_year, "\\$")) %>% 
  mutate(food_costs_per_year = str_remove(food_costs_per_year, ","))
```

```{r}
#convert chr to num
dog_rankings_c <- dog_rankings_c %>% 
  mutate(lifetime_cost = as.numeric(lifetime_cost)) %>% 
  mutate(intelligence_percent = as.numeric(intelligence_percent)) %>% 
  mutate(purchase_price = as.numeric(purchase_price)) %>% 
  mutate(food_costs_per_year = as.numeric(food_costs_per_year))

glimpse(dog_rankings_c)
```
```{r}
write.csv(dog_rankings_c, "datasets/dog_rankings_c", row.names = F)
```

## Fiona Data Exploration
```{r}
#popularity by type of dog
dog_rankings_c %>% 
  group_by(type) %>% 
  summarize(popularity = mean(popularity_ranking)) %>% 
  arrange(desc(popularity))
```

Two type of scoring for family/kid friendly

1. score_for_kids: A rating indicating how suitable the breed is for families with children. Higher scores suggest greater suitability. Intelligence + longevity + ailments + cost scores (50% each) + 100% grooming score. highest possible score: 5

2. suitability_for_children: The suitability of the breed for families with children, typically measured on a scale (1 = high suitability, 2= medium suitability, 3= low suitability).
```{r}
#good for kids
dog_rankings_c %>% 
  arrange(suitability_for_children) %>% 
  select(breed, suitability_for_children, score_for_kids) %>% 
  filter(score_for_kids >= 4.00) %>% 
  arrange(-score_for_kids)
```
```{r}
#most expensive purchase dog
dog_rankings_c %>% 
  select(breed, type, lifetime_cost, purchase_price, food_costs_per_year, grooming_frequncy) %>% 
  slice_max(purchase_price, n=20) %>% 
  ggplot(aes(x=reorder(breed, purchase_price), y=purchase_price))+
  geom_col()+
  coord_flip()
```
```{r}
#cheapest purchase dog
dog_rankings_c %>% 
  select(breed, type, lifetime_cost, purchase_price, food_costs_per_year, grooming_frequncy) %>% 
  slice_min(purchase_price, n=20) %>% 
  ggplot(aes(x=reorder(breed, -purchase_price), y=purchase_price))+
  geom_col()+
  coord_flip()
```

```{r}
#most common breed in shelter
shelter_dogs %>% 
  count(breed) %>% 
  arrange(desc(n))
```

## Allyson Data Exploration
```{r}
dog_rankings_c
```
```{r}
#Graphs for the Presentation
  # 1) Is popularity determined by breed intelligence?
dog_rankings_c %>% 
  ggplot(aes(x=popularity_ranking, y=intelligence_rank))+
  geom_col()

#No correlation, popular dogs have no correlation with intelligence
```
```{r}
# 2) Does breed size (smaller or larger dogs) correlate with number of genetic aliments?
dog_rankings_c %>% 
  ggplot(aes(x=size_description, y=number_of_genetic_ailments))+
  geom_boxplot()

#No correlation
```

```{r}
# 3) Is there a correlation between size and longevity? Do smaller dogs live longer according to this data?
dog_rankings_c %>% 
  ggplot(aes(x=size_description, y=longevity_years))+
  geom_col()

#There is a correlation! Small dogs live longer!
```

```{r}
# 4) Do small dogs tend to be more popular?
dog_rankings_c %>% 
  ggplot(aes(x=popularity_ranking, y=size_description))+
  geom_col()
```

## Jenna Data Exploration
```{r}
summary(shelter_dogs)
```

```{r}
#number of male and female dogs
shelter_dogs %>% 
  select(sex, breed) %>% 
  count(sex)
```

```{r}
#breeds in the shelter
shelter_dogs %>% 
  filter(breed != "Adoptable From:", breed != "Unknown Mix") %>% 
  count(breed) %>% 
  arrange(-n)
```

```{r}
#different breeds of SMALL dogs in the shelter waiting to be adopted
small_breeds <- shelter_dogs %>% 
  select(breed, size) %>% 
  filter(breed != "Unknown Mix", size == "small") %>% 
  count(breed) %>% 
  arrange(-n)
small_breeds
```

```{r}
#different breeds of MEDIUM dogs in the shelter waiting to be adopted
medium_breeds <- shelter_dogs %>% 
  select(breed, size) %>% 
  filter(breed != "Unknown Mix", size == "medium") %>% 
  count(breed) %>% 
  arrange(-n)
medium_breeds
```

```{r}
#different breeds of LARGE dogs in the shelter waiting to be adopted
large_breeds <- shelter_dogs %>% 
  select(breed, size) %>% 
  filter(breed != "Unknown Mix", breed != "Adoptable From:", size == "large") %>% 
  count(breed) %>% 
  arrange(-n)
large_breeds
```

```{r}
dog_rankings %>% 
  count(type)
```

```{r}
#age distribution
shelter_dogs %>% 
  ggplot(aes(x = age))+
  geom_density(color = 'black', fill = 'hotpink', alpha = 0.6)
```


##Notable Graphs for Presentation Visuals

```{r}
#theme and colors
colors <- paletteer::palettes_d_names
my_palette <- paletteer_d("calecopal::superbloom1")
barplot(rep(1,6), axes=FALSE, col=my_palette)
```

```{r}
#size and longevity correlation
dog_rankings_c %>% 
  ggplot(aes(x=size_description, y=longevity_years, fill = size_description))+
  geom_boxplot()+
  labs(title = "Correlation of Size and Longevity", 
       x = "Dog Size", 
       y = "Longevity (years)",
       fill = "Size")+
  theme_few()+
  scale_fill_manual(values=my_palette)
```

```{r}
#age distribution
shelter_dogs %>% 
  ggplot(aes(x = age))+
  geom_density(color = 'black', fill = 'lightblue', alpha = 0.6)+
  labs(title = "Distribution of Shelter Dog's Age", 
       x = "Age", 
       y = "Density")+
  theme_few()+
  scale_fill_manual(values=my_palette)
```

```{r}
#most common breed in shelter
shelter_dogs %>%
  count(breed) %>%
  slice_max(n, n = 5) %>%
  ggplot(aes(x = reorder(breed, n), y = n, fill = breed)) +
  geom_bar(stat = "identity")+
  coord_flip()+
  labs(title = "Most Common Breeds in Shelter", 
       x= "Breed", 
       y= "Count", 
       fill = "Breed")+
  theme_few()+
  scale_fill_manual(values=my_palette)
```
```{r}
#install.packages("colourpicker")
#install.packages("magick")
#install.packages("scales")
#install.packages("imager")
library(magick)
library(scales)
library(imager)
library(colourpicker)
img <- image_read("colors.png")

# Convert to a data frame
img_data <- as.data.frame(img, wide = "c") 

# Extract unique colors and select top 6 most frequent
top_colors <- img_data %>%
  count(rgb = rgb(c.1, c.2, c.3, maxColorValue = 255)) %>%
  arrange(desc(n)) %>%
  head(6)  # Select top 6 colors

# Display the colors
show_col(top_colors$rgb)
```


```{r}
my_extended_palette <- c("#FDE725", "#7AD151", "#22A884", "#2A788E", "#414487","#FBAE17", "#468C48", "#3F4DA2")

dog_rankings_c %>%
  ggplot(aes(x = type, y = popularity_ranking, fill = type))+
  geom_boxplot()+
  coord_flip()+
  labs(title = "Most Popular Type of Dog", 
       x= "Dog Type", 
       y= "Popularity Score", 
       fill = "Dog Type")+
  theme_few()+
  scale_fill_manual(values=my_extended_palette)
```


##Shiny App 

#Quiz Question #1: What type of dog are you looking for?
  Answer choices would include: Herding, hound, non-sporting, sporting, terrier, toy, or working
  
#Quiz Question #2

```{r}
library(shiny)

ui <- fluidPage(
  title = "Adopt don't Shop!",
  sidebarPanel()
  
)

server <- function(input, output, session) {
  
}

shinyApp(ui, server)
```


```{r}
install.packages("shinythemes")
```

```{r}
install.packages("DT")

```

```{r}

```


##Shelter Dogs Shiny App - Allyson Hafezi

```{r}
library(shiny)
library(shinythemes)
library(tidyverse)

# Load dataset
dog_data <- read_csv("ShelterDogs.csv")

ui <- fluidPage(
  theme = shinytheme("cerulean"),
  tags$head(tags$style(HTML("
    body { background-color: #f0f8ff; }
    .title-panel { text-align: center; padding: 20px; }
    .panel-box {
      background: #ffffff;
      padding: 20px;
      border-radius: 10px;
      box-shadow: 0 2px 10px rgba(0,0,0,0.1);
      margin: 20px auto;
      max-width: 800px;
    }
    h1, h3, label { color: #333333; }
    .btn-primary { background-color: #ff6b6b; border-color: #ff6b6b; }
    .btn-primary:hover { background-color: #ff4757; border-color: #ff4757; }
    .dog-card {
      background: #fff;
      padding: 15px;
      margin-bottom: 15px;
      border-radius: 8px;
      box-shadow: 0 2px 5px rgba(0,0,0,0.1);
    }
  "))),
  div(class = "title-panel", h1("ðŸ¶ Find Your Perfect Dog! ðŸ¾")),
 
  # Row for all questions (displayed on one panel at the top)
  fluidRow(
    column(12,
           div(class = "panel-box",
               h3("Tell us about your ideal dog!"),
               selectInput("size", "Preferred Dog Size:",
                           choices = c("Not Important", unique(dog_data$size))),
               selectInput("coat", "Preferred Coat Type:",
                           choices = c("Not Important", unique(dog_data$coat))),
               selectInput("coat_color", "What coat color do you prefer?",
                           choices = c("Not Important", unique(dog_data$color))),
               selectInput("sex", "Preferred Sex:",
                           choices = c("Not Important", unique(dog_data$sex))),
               sliderInput("age", "Preferred Age Range:",
                           min = floor(min(dog_data$age, na.rm = TRUE)),
                           max = ceiling(max(dog_data$age, na.rm = TRUE)),
                           value = c(floor(min(dog_data$age, na.rm = TRUE)),
                                     ceiling(max(dog_data$age, na.rm = TRUE))),
                           step = 1),
               selectInput("good_with_kids", "Good with children?",
                           choices = c("Not Important", "yes")),
               selectInput("good_with_males", "Good with male dogs?",
                           choices = c("Not Important", "yes")),
               selectInput("good_with_females", "Good with female dogs?",
                           choices = c("Not Important", "yes")),
               selectInput("good_with_cats", "Good with cats?",
                           choices = c("Not Important", "yes")),
               selectInput("house_trained", "House Trained?",
                           choices = c("Not Important", "yes")),
               selectInput("neutered", "Neutered?",
                           choices = c("Not Important", "yes", "no")),
               br(),
               actionButton("find", "ðŸ¾ Find My Dog!", class = "btn-primary btn-lg")
           )
    )
  ),
 
  # Row for the results (displayed below the questions)
  fluidRow(
    column(12,
           div(class = "panel-box",
               h3("Your Best Matches:"),
               h4(textOutput("num_results")),
               uiOutput("results")
           )
    )
  )
)

server <- function(input, output, session) {
 
  filtered_data <- eventReactive(input$find, {
    df <- dog_data
    if (input$size != "Not Important") {
      df <- df %>% filter(size == input$size)
    }
    if (input$coat != "Not Important") {
      df <- df %>% filter(coat == input$coat)
    }
    if (input$coat_color != "Not Important") {
      df <- df %>% filter(color == input$coat_color)
    }
    if (input$sex != "Not Important") {
      df <- df %>% filter(sex == input$sex)
    }
    df <- df %>% filter(age >= input$age[1] & age <= input$age[2])
    if (input$good_with_kids == "yes") {
      df <- df %>% filter(likes_children == "yes")
    }
    if (input$good_with_males == "yes") {
      df <- df %>% filter(get_along_males == "yes")
    }
    if (input$good_with_females == "yes") {
      df <- df %>% filter(get_along_females == "yes")
    }
    if (input$good_with_cats == "yes") {
      df <- df %>% filter(get_along_cats == "yes")
    }
    if (input$house_trained == "yes") {
      df <- df %>% filter(housebroken == "yes")
    }
    if (input$neutered != "Not Important") {
      df <- df %>% filter(neutered == input$neutered)
    }
    df
  })
 
  output$num_results <- renderText({
    paste("Number of Matches:", nrow(filtered_data()))
  })
 
  output$results <- renderUI({
    df <- filtered_data()
    if (nrow(df) == 0) {
      return(h4("No matches found! Try adjusting your preferences.",
                style = "color: red; text-align: center;"))
    }
    lapply(1:nrow(df), function(i) {
      div(class = "dog-card",
          h4(df$name[i], style = "color: #ff6b6b;"),
          p(strong("Breed:"), df$breed[i]),
          p(strong("Sex:"), df$sex[i]),
          p(strong("Size:"), df$size[i]),
          p(strong("Age:"), df$age[i], "years"),
          p(strong("Coat:"), df$coat[i]),
          p(strong("Coat Color:"), df$color[i])
      )
    })
  })
}

shinyApp(ui = ui, server = server)
```







```{r}
library(shiny)


# UI
ui <- fluidPage(
  theme = shinytheme("superhero"),
  div(class = "title-panel", h1("ðŸ¶ Find Your Perfect Dog! ðŸ¾", style = "color: #ff6b6b;")),
  sidebarLayout(
    sidebarPanel(
      class = "sidebar-panel",
      h3("Tell us about your ideal dog!"),
      selectInput("size", "Preferred Dog Size:", choices = c("Any", unique(dog_data$size))),
      selectInput("coat", "Preferred Coat Type:", choices = c("Any", unique(dog_data$coat))),
      sliderInput("age", "Preferred Age Range:",
                  min = floor(min(dog_data$age, na.rm = TRUE)),
                  max = ceiling(max(dog_data$age, na.rm = TRUE)),
                  value = c(floor(min(dog_data$age, na.rm = TRUE)), ceiling(max(dog_data$age, na.rm = TRUE))),
                  step = 1),
      selectInput("activity", "Activity Level:", choices = c("Any", "Low", "Medium", "High")),
      selectInput("experience", "Dog Owner Experience:", choices = c("Any", "Beginner", "Intermediate", "Experienced")),
      selectInput("good_with_kids", "Good with children?", choices = c("Any", "yes")),
      selectInput("good_with_males", "Good with male dogs?", choices = c("Any", "yes")),
      selectInput("good_with_females", "Good with female dogs?", choices = c("Any", "yes")),
      selectInput("good_with_cats", "Good with cats?", choices = c("Any", "yes")),
      selectInput("house_trained", "House Trained?", choices = c("Any", "yes")),
      selectInput("neutered", "Neutered?", choices = c("Any", "yes", "no")),
      br(),
      actionButton("find", "ðŸ¾ Find My Dog!", class = "btn-primary btn-lg")
    ),
    mainPanel(
      class = "main-panel",
      h3("ðŸ• Your Best Matches:"),
      h4(textOutput("num_results"), style = "color: #333333; text-align: center;"),
      uiOutput("results"),
      h3("ðŸ¾ Dog Rankings:", style = "margin-top: 30px;"),
      dataTableOutput("rankings_table")
    )
  )
)

# Server
server <- function(input, output, session) {
  filtered_data <- eventReactive(input$find, {
    df <- dog_data
    if (input$size != "Any") df <- df[df$size == input$size,]
    if (input$coat != "Any") df <- df[df$coat == input$coat,]
    df <- df[df$age >= input$age[1] & df$age <= input$age[2],]
    if (input$good_with_kids == "yes") df <- df[df$likes_children == "yes",]
    if (input$good_with_males == "yes") df <- df[df$get_along_males == "yes",]
    if (input$good_with_females == "yes") df <- df[df$get_along_females == "yes",]
    if (input$good_with_cats == "yes") df <- df[df$get_along_cats == "yes",]
    if (input$house_trained == "yes") df <- df[df$housebroken == "yes",]
    if (input$neutered != "Any") df <- df[df$neutered == input$neutered,]
    df
  })
  
  output$num_results <- renderText({
    paste("Number of Matches:", nrow(filtered_data()))
  })
  
  output$results <- renderUI({
    dog_list <- filtered_data()
    if (nrow(dog_list) == 0) {
      return(h4("No matches found! Try adjusting your preferences.", style = "color: red; text-align: center;"))
    }
    dog_ui <- lapply(1:nrow(dog_list), function(i) {
      dog <- dog_list[i, ]
      div(class = "dog-card",
          h4(dog$name, style = "color: #ff6b6b;"),
          p(strong("Breed:"), dog$breed),
          p(strong("Size:"), dog$size),
          p(strong("Age:"), dog$age, "years"),
          p(strong("Coat:"), dog$coat),
          p(strong("Activity Level:"), dog$activity_level)
      )
    })
    do.call(tagList, dog_ui)
  })
  
  output$rankings_table <- renderDataTable({
    datatable(dog_rankings, options = list(pageLength = 10, autoWidth = TRUE))
  })
}

shinyApp(ui, server)
```



